---
title: "[07] Conditional logic examples Guide"
output:
  md_document:
    variant: gfm
output_dir: ../workshops
knitr:
  opts_knit:
    root.dir: ..
---

This tutorial complements `07_architecting_conditional_logic_examples.R` and unpacks the workshop on conditional logic examples. You will see how it advances the Architecting Systems sequence while building confidence with base R and tidyverse tooling.

## Setup

- Ensure you have opened the `archr` project root (or set your working directory there) before running any code.
- Open the workshop script in RStudio so you can execute lines interactively with `Ctrl+Enter` or `Cmd+Enter`.
- Create a fresh R session to avoid conflicts with leftover objects from earlier workshops.

## Skills

- Navigate the script `07_architecting_conditional_logic_examples.R` within the Architecting Systems module.
- Connect the topic "Conditional logic examples" to systems architecting decisions.
- Load packages with `library()` and verify they attach without warnings.
- Chain tidyverse verbs with `%>%` to explore stakeholder or architecture tables.
- Define custom functions to package repeatable logic.

## Application

### Step 1 – Load Packages

SETUP #################################### Load packages.

```{r step_01, eval=FALSE}
library(dplyr)
library(archr)
library(ggplot2)
```

### Step 2 – Create `a`

Make architectures to work with.

```{r step_02, eval=FALSE}
a = enumerate_sf(n = c(2,2,2))
```

### Step 3 – Define `get_cost()`

Suppose. Create the helper function `get_cost()` so you can reuse it throughout the workshop.

```{r step_03, eval=FALSE}
get_cost = function(d1, d2, d3){
  # Let's make some objects JUST for testing - be sure to comment them out when done
  # d1 = 1; d2 = 1; d3 = 1
```

### Step 4 – Create `m1`

When d1 is 1, make m1 3; When d1 is 0, make m1 0 and repeat for other problem Or, try using case when.

```{r step_04, eval=FALSE}
  m1 = case_when(d1 == 1 ~ 3, d1 == 0 ~ 0)
  m2 = case_when(d2 == 1 ~ 6, d2 == 0 ~ 0)
  m3 = case_when(d3 == 1 ~ 9, d3 == 0 ~ 0)
```

### Step 5 – Create `metric`

Compute overall metric. Create the object `metric` so you can reuse it in later steps.

```{r step_05, eval=FALSE}
  metric = m1+m2+m3
  return(metric)
}
```

### Step 6 – Practice the Pipe

Let's try it. Use the `%>%` operator to pass each result to the next tidyverse verb.

```{r step_06, eval=FALSE}
a %>%
  mutate(cost = get_cost(d1,d2,d3))
```

### Step 7 – Practice the Pipe

Use both. Use the `%>%` operator to pass each result to the next tidyverse verb.

```{r step_07, eval=FALSE}
data = a %>%
  mutate(cost = get_cost(d1,d2,d3))
```

### Step 8 – Practice the Pipe

View histogram of cost values for your architectures!

```{r step_08, eval=FALSE}
data$cost %>% hist()
```

### Step 9 – Clear Objects

Cleanup! Remove objects from the environment to prevent name clashes.

```{r step_09, eval=FALSE}
rm(list = ls())
```

## Learning Checks

**Learning Check 1.** How do you run the entire workshop script after you have stepped through each section interactively?

<details>
<summary>Show answer</summary>

Use `source(file.path("workshops", "07_architecting_conditional_logic_examples.R"))` from the Console or press the Source button while the script is active.

</details>

**Learning Check 2.** Why does the script begin by installing or loading packages before exploring the exercises?

<details>
<summary>Show answer</summary>

Those commands make sure the required libraries are available so every subsequent code chunk runs without missing-function errors.

</details>

**Learning Check 3.** How does the `%>%` pipeline help you reason about multi-step transformations in this script?

<details>
<summary>Show answer</summary>

It keeps each operation in sequence without creating temporary variables, so you can narrate the data story line by line.

</details>

**Learning Check 4.** How can you build confidence that a newly defined function behaves as intended?

<details>
<summary>Show answer</summary>

Call it with the sample input from the script, examine the output, then try a new input to see how the behaviour changes.

</details>
