---
title: "[13] Minimum sample sizing utility Guide"
output:
  md_document:
    variant: gfm
output_dir: ../workshops
knitr:
  opts_knit:
    root.dir: ..
---

This tutorial complements `13_enumeration_sample_sizing.R` and unpacks the workshop on minimum sample sizing utility. You will see how it advances the Enumeration sequence while building confidence with base R and tidyverse tooling.

## Setup

- Ensure you have opened the `archr` project root (or set your working directory there) before running any code.
- Open the workshop script in RStudio so you can execute lines interactively with `Ctrl+Enter` or `Cmd+Enter`.
- Create a fresh R session to avoid conflicts with leftover objects from earlier workshops.

## Skills

- Navigate the script `13_enumeration_sample_sizing.R` within the Enumeration module.
- Connect the topic "Minimum sample sizing utility" to systems architecting decisions.
- Load packages with `library()` and verify they attach without warnings.
- Chain tidyverse verbs with `%>%` to explore stakeholder or architecture tables.
- Define custom functions to package repeatable logic.

## Process Overview









```mermaid
flowchart LR
    A[Define get_minsize()] --> B[Load Packages]
    B[Load Packages] --> C[Run the Code Block]
    C[Run the Code Block] --> D[Clear Objects]
```

## Application

### Step 1 – Define `get_minsize()`

Create the helper function `get_minsize()` so you can reuse it throughout the workshop.

```{r step_01, eval=FALSE}
get_minsize = function(data, decisions = c("d1")){
  library(dplyr)
  library(tidyr)
```

### Step 2 – Practice the Pipe

Use the `%>%` operator to pass each result to the next tidyverse verb.

```{r step_02, eval=FALSE}
  data %>%
    select(any_of(decisions)) %>%
    # Count up k, the number of unique alternatives per decision
    summarize(across(everything(), .fns = ~length(unique(.x)))) %>%
    # Pivot it longer
    tidyr::pivot_longer(
      cols = everything(), 
      names_to = "k", values_to = "value") %>%
    # Count up the minimum number of architectures necessary to test
    # that number of main effects
    summarize(minsize = 1 + sum(value - 1)) %>%
    # Return it as a vector
    with(minsize)        
}
```

### Step 3 – Load Packages

Attach dplyr to make its functions available.

```{r step_03, eval=FALSE}
library(dplyr)
library(tidyr)
library(archr)
```

### Step 4 – Create `a`

Create the object `a` so you can reuse it in later steps.

```{r step_04, eval=FALSE}
a = enumerate_sf(n = c(2,5,4))
```

### Step 5 – Run the Code Block

to count sample size needed...

```{r step_05, eval=FALSE}
get_minsize(a, decisions = c("d1"))
get_minsize(a, decisions = c("d1", "d2"))
get_minsize(a, decisions = c("d1", "d2", "d3"))
```

### Step 6 – Clear Objects

Remove objects from the environment to prevent name clashes.

```{r step_06, eval=FALSE}
rm(list = ls())
```

## Learning Checks











**Learning Check 1.** Which libraries does Step 1 attach, and why do you run that chunk before others?

<details>
<summary>Show answer</summary>

It attaches dplyr and tidyr, ensuring their functions are available before you execute the downstream code.

</details>

**Learning Check 2.** After Step 2, what does `cols` capture?

<details>
<summary>Show answer</summary>

It creates `cols` that computes summary statistics, and selects a focused set of columns. Use the `%>%` operator to pass each result to the next tidyverse verb.

</details>

**Learning Check 3.** After Step 4, what does `a` capture?

<details>
<summary>Show answer</summary>

It creates `a` that enumerates architecture combinations with `archr` helpers. Create the object `a` so you can reuse it in later steps.

</details>

**Learning Check 4.** Why is Step 5 (‘Run the Code Block’) necessary in this guide?

<details>
<summary>Show answer</summary>

to count sample size needed...

</details>
