---
title: "[20] Custom metric development Guide"
output:
  md_document:
    variant: gfm
output_dir: ../workshops
knitr:
  opts_knit:
    root.dir: ..
---

This tutorial complements `20_enumeration_custom_metrics.R` and unpacks the workshop on custom metric development. You will see how it advances the Enumeration sequence while building confidence with base R and tidyverse tooling.

## Setup

- Ensure you have opened the `archr` project root (or set your working directory there) before running any code.
- Open the workshop script in RStudio so you can execute lines interactively with `Ctrl+Enter` or `Cmd+Enter`.
- Create a fresh R session to avoid conflicts with leftover objects from earlier workshops.

## Skills

- Navigate the script `20_enumeration_custom_metrics.R` within the Enumeration module.
- Connect the topic "Custom metric development" to systems architecting decisions.
- Load packages with `library()` and verify they attach without warnings.
- Chain tidyverse verbs with `%>%` to explore stakeholder or architecture tables.
- Define custom functions to package repeatable logic.

## Process Overview









```mermaid
flowchart LR
    A[Load Packages] --> B[Run the Code Block]
    B[Run the Code Block] --> C[Run the Code Block (Step 7)]
    C[Run the Code Block (Step 7)] --> D[Clear Objects]
```

## Application

### Step 1 – Load Packages

Attach dplyr to make its functions available.

```{r step_01, eval=FALSE}
library(dplyr)
library(archr)
```

### Step 2 – Define `get_percentages()`

Can be hard to compare visually, but let's check them numerically...

```{r step_02, eval=FALSE}
get_percentages = function(x, did = "d2"){
  # Testing values
  #x = mini$d2
```

### Step 3 – Practice the Pipe

Get the tally per alternative.

```{r step_03, eval=FALSE}
  tally = tibble(altid = x) %>%
    group_by(altid) %>%
    summarize(count = n())
  # Calculate percentages
  percentages = tally %>% 
    mutate(total = sum(count)) %>%
    mutate(percent = count / total) %>%
    mutate(label = round(percent * 100, 1)) %>%
    mutate(label = paste0(label, "%"))
  # Add a decision id and format
  percentages = percentages %>% 
    mutate(did = did) %>%
    select(did, altid, count, total, percent, label)
```

### Step 4 – Run the Code Block

Execute the block and pay attention to the output it produces.

```{r step_04, eval=FALSE}
  return(percentages)
}
```

### Step 5 – Create `a`

Make an architecture matrix of 3 binary decisions.

```{r step_05, eval=FALSE}
a = enumerate_sf(n = c(2,2,2))
```

### Step 6 – Practice the Pipe

Repeat this code several times.

```{r step_06, eval=FALSE}
sample = a %>%
  group_by(d1) %>%
  sample_n(size = 2)
```

### Step 7 – Run the Code Block

Execute the block and pay attention to the output it produces.

```{r step_07, eval=FALSE}
get_percentages(x = sample$d1, did = "d1")
get_percentages(x = sample$d2, did = "d2")
get_percentages(x = sample$d3, did = "d3")
```

### Step 8 – Practice the Pipe

One way to improve balance is to stratify by MULTIPLE variables. (though NEVER by ALL variables).

```{r step_08, eval=FALSE}
sample = a %>%
  group_by(d1, d2) %>%
  sample_n(size = 2)
```

### Step 9 – Run the Code Block

Execute the block and pay attention to the output it produces.

```{r step_09, eval=FALSE}
get_percentages(x = sample$d1, did = "d1")
get_percentages(x = sample$d2, did = "d2")
get_percentages(x = sample$d3, did = "d3")
```

### Step 10 – Clear Objects

A little better! Remove objects from the environment to prevent name clashes.

```{r step_10, eval=FALSE}
rm(list = ls())
```

## Learning Checks











**Learning Check 1.** What role does the helper `get_percentages()` defined in Step 2 play in this workflow?

<details>
<summary>Show answer</summary>

It packages reusable logic needed by later steps.

</details>

**Learning Check 2.** Which libraries does Step 1 attach, and why do you run that chunk before others?

<details>
<summary>Show answer</summary>

It attaches dplyr and archr, ensuring their functions are available before you execute the downstream code.

</details>

**Learning Check 3.** After Step 3, what does `tally` capture?

<details>
<summary>Show answer</summary>

It creates `tally` that groups the data by relevant variables, and computes summary statistics. Get the tally per alternative.

</details>

**Learning Check 4.** After Step 5, what does `a` capture?

<details>
<summary>Show answer</summary>

It creates `a` that enumerates architecture combinations with `archr` helpers. Make an architecture matrix of 3 binary decisions.

</details>
