---
title: "[27] Utility function design Guide"
output:
  md_document:
    variant: gfm
output_dir: ../workshops
knitr:
  opts_knit:
    root.dir: ..
---

This tutorial complements `27_evaluation_utility_functions.R` and unpacks the workshop on utility function design. You will see how it advances the Evaluation sequence while building confidence with base R and tidyverse tooling.

## Setup

- Ensure you have opened the `archr` project root (or set your working directory there) before running any code.
- Open the workshop script in RStudio so you can execute lines interactively with `Ctrl+Enter` or `Cmd+Enter`.
- Create a fresh R session to avoid conflicts with leftover objects from earlier workshops.

## Skills

- Navigate the script `27_evaluation_utility_functions.R` within the Evaluation module.
- Connect the topic "Utility function design" to systems architecting decisions.
- Load packages with `library()` and verify they attach without warnings.

## Process Overview









```mermaid
flowchart LR
    A[Load Packages] --> B[Create u_pay]
    B[Create u_pay] --> C[Create w_e]
    C[Create w_e] --> D[Clear Objects]
```

## Application

### Step 1 – Load Packages

Setup #################################################. Attach dplyr to make its functions available.

```{r step_01, eval=FALSE}
library(dplyr)
library(ggplot2)
```

### Step 2 – Create `p_fail`

Define probabilities. Create the object `p_fail` so you can reuse it in later steps.

```{r step_02, eval=FALSE}
p_fail = 0.05
p_works = 1 - p_fail
```

### Step 3 – Create `pay_min`

Define utility function for pay Choose what means min utility (0) for you.

```{r step_03, eval=FALSE}
pay_min = 0
# Choose what means max utility (1) for you
pay_max = 10000
```

### Step 4 – Create `pay_works`

Calculate utility of drone working (by rescaling).

```{r step_04, eval=FALSE}
pay_works = 10000
u_pay_works = (pay_works - pay_min) / (pay_max - pay_min)
```

### Step 5 – Create `pay_fail`

Calculate utility if drone fails.

```{r step_05, eval=FALSE}
pay_fail = -5000
u_pay_fail = (pay_fail - pay_min) / (pay_max - pay_min)
```

### Step 6 – Create `u_pay`

Calculate expected utility that you would earn Expected utility here is ~0.93.

```{r step_06, eval=FALSE}
u_pay = p_works * u_pay_works + p_fail * u_pay_fail
u_pay # View
```

### Step 7 – Create `e_min`

Define utility function for emissions.

```{r step_07, eval=FALSE}
e_min = 0
e_max = 20
```

### Step 8 – Create `e_works`

Create the object `e_works` so you can reuse it in later steps.

```{r step_08, eval=FALSE}
e_works = 10
```

### Step 9 – Create `u_e_works`

Create the object `u_e_works` so you can reuse it in later steps.

```{r step_09, eval=FALSE}
u_e_works = (e_works - e_min) / (e_max - e_min)
```

### Step 10 – Create `e_fail`

Create the object `e_fail` so you can reuse it in later steps.

```{r step_10, eval=FALSE}
e_fail = 3
u_e_fail = (e_fail - e_min) / (e_max - e_min)
```

### Step 11 – Create `u_e`

So the expected utility here is ~0.52 on a scale from 0 to 20. Calculate expected utility for emissions.

```{r step_11, eval=FALSE}
u_e = p_works * u_e_works + p_fail * u_e_fail
u_e # View
1 - u_e
```

### Step 12 – Create `w_e`

Define weights. Create the object `w_e` so you can reuse it in later steps.

```{r step_12, eval=FALSE}
w_e = 0.30
w_pay = 0.70
```

### Step 13 – Create `t`

Combine utilities. Create the object `t` so you can reuse it in later steps.

```{r step_13, eval=FALSE}
t = tibble(
  type = c("emissions", "pay"),           # Types of utilities
  # remember to make high emissions BAD
  # remember direction matters!
  u = c(1 - u_e, u_pay),                      # Utility values
  w = c(w_e, w_pay)                       # Weights
)
```

### Step 14 – Run the Code Block

Execute the block and pay attention to the output it produces.

```{r step_14, eval=FALSE}
t
```

### Step 15 – Create `combined_additive`

Calculate combined utility additively.

```{r step_15, eval=FALSE}
combined_additive = sum(t$u * t$w)
combined_additive
```

### Step 16 – Create `combined_multiplicative`

Calculate combined utility multiplicatively.

```{r step_16, eval=FALSE}
combined_multiplicative = prod(t$u * t$w)
combined_multiplicative
```

### Step 17 – Clear Objects

Remove objects from the environment to prevent name clashes.

```{r step_17, eval=FALSE}
rm(list = ls())
```

## Learning Checks











**Learning Check 1.** Which libraries does Step 1 attach, and why do you run that chunk before others?

<details>
<summary>Show answer</summary>

It attaches dplyr and ggplot2, ensuring their functions are available before you execute the downstream code.

</details>

**Learning Check 2.** After Step 2, what does `p_fail` capture?

<details>
<summary>Show answer</summary>

It creates `p_fail`. Define probabilities. Create the object `p_fail` so you can reuse it in later steps.

</details>

**Learning Check 3.** After Step 3, what does `pay_min` capture?

<details>
<summary>Show answer</summary>

It creates `pay_min`. Define utility function for pay Choose what means min utility (0) for you.

</details>

**Learning Check 4.** After Step 4, what does `pay_works` capture?

<details>
<summary>Show answer</summary>

It creates `pay_works`. Calculate utility of drone working (by rescaling).

</details>
