---
title: "[03] First R script practice Guide"
output:
  md_document:
    variant: gfm
output_dir: ../workshops
knitr:
  opts_knit:
    root.dir: ..
---

This tutorial complements `03_stakeholder_first_r_script.R` and unpacks the workshop on first r script practice. You will see how it advances the Stakeholder Analysis sequence while building confidence with base R and tidyverse tooling.

## Setup

- Ensure you have opened the `archr` project root (or set your working directory there) before running any code.
- Open the workshop script in RStudio so you can execute lines interactively with `Ctrl+Enter` or `Cmd+Enter`.
- Create a fresh R session to avoid conflicts with leftover objects from earlier workshops.

## Skills

- Navigate the script `03_stakeholder_first_r_script.R` within the Stakeholder Analysis module.
- Connect the topic "First R script practice" to systems architecting decisions.
- Document observations from running each code block in your lab notebook.

## Application

### Step 1 – Try Arithmetic

My first script. Evaluate a quick arithmetic expression to observe R's console output.

```{r step_01, eval=FALSE}
1 + 2 
x = 2
x
x + 2
c(2,3,4)
```

### Step 2 – Create `myvec`

Create the object `myvec` so you can reuse it in later steps.

```{r step_02, eval=FALSE}
myvec = c(1,2,3,4)
myvec
```

## Learning Checks

**Learning Check 1.** How do you run the entire workshop script after you have stepped through each section interactively?

<details>
<summary>Show answer</summary>

Use `source(file.path("workshops", "03_stakeholder_first_r_script.R"))` from the Console or press the Source button while the script is active.

</details>

**Learning Check 2.** In your own words, what key idea does the topic "First R script practice" reinforce?

<details>
<summary>Show answer</summary>

It highlights how first r script practice supports the overall systems architecting process in this course.

</details>

**Learning Check 3.** Where should you look if you need to reset your setup before re-running the script?

<details>
<summary>Show answer</summary>

Return to the Setup section to confirm your working directory, packages, and clean R session.

</details>
