---
title: "[09] Cost function modeling examples Guide"
output:
  md_document:
    variant: gfm
output_dir: ../workshops
knitr:
  opts_knit:
    root.dir: ..
---

This tutorial complements `09_architecting_cost_function_examples.R` and unpacks the workshop on cost function modeling examples. You will see how it advances the Architecting Systems sequence while building confidence with base R and tidyverse tooling.

## Setup

- Ensure you have opened the `archr` project root (or set your working directory there) before running any code.
- Open the workshop script in RStudio so you can execute lines interactively with `Ctrl+Enter` or `Cmd+Enter`.
- Create a fresh R session to avoid conflicts with leftover objects from earlier workshops.

## Skills

- Navigate the script `09_architecting_cost_function_examples.R` within the Architecting Systems module.
- Connect the topic "Cost function modeling examples" to systems architecting decisions.
- Load packages with `library()` and verify they attach without warnings.
- Chain tidyverse verbs with `%>%` to explore stakeholder or architecture tables.
- Define custom functions to package repeatable logic.
- Iterate on visualisations built with `ggplot2`.

## Application

### Step 1 – Load Packages

Setup #####################################################. Attach dplyr to make its functions available.

```{r step_01, eval=FALSE}
library(dplyr)
library(ggplot2)
```

### Step 2 – Create `data`

smartboards. Create the object `data` so you can reuse it in later steps.

```{r step_02, eval=FALSE}
data = tribble(
  ~unit, ~cost,
  1,     2000,
  10,    16000,
  50,    90000,
  100,   140000,
)
```

### Step 3 – Start a ggplot

Initialize a ggplot so you can layer geoms and customise aesthetics.

```{r step_03, eval=FALSE}
ggplot() +
  geom_point(data = data, mapping = aes(x = unit, y = cost))
```

### Step 4 – Practice the Pipe

Estimate a linear model of cost.

```{r step_04, eval=FALSE}
m = data %>% lm(formula = cost ~ unit)
# Cost = $4,920 + $1,418 * unit
```

### Step 5 – Define `get_cost()`

Equivalently stated as a function:.

```{r step_05, eval=FALSE}
get_cost = function(unit){ 4920 + 1418 * unit }
```

### Step 6 – Create `data2`

Calculate it. Create the object `data2` so you can reuse it in later steps.

```{r step_06, eval=FALSE}
data2 = tibble(
  unit = 150,
  # pred = get_cost(unit), # this is equivalent
  pred = predict(m, newdata = tibble(unit))
)
```

### Step 7 – Create `data3`

Create the object `data3` so you can reuse it in later steps.

```{r step_07, eval=FALSE}
data3 = tibble(
  unit = 1:150,
  pred = predict(m, newdata = tibble(unit))
)
```

### Step 8 – Run the Code Block

Execute the block and pay attention to the output it produces.

```{r step_08, eval=FALSE}
data3
```

### Step 9 – Start a ggplot

Let's visualize them. Initialize a ggplot so you can layer geoms and customise aesthetics.

```{r step_09, eval=FALSE}
ggplot() +
  geom_point(data = data, mapping = aes(x = unit, y = cost)) +
  geom_line(data = data3, mapping = aes(x = unit, y = pred)) +
  geom_point(data = data2, mapping = aes(x = unit, y = pred), color = "red")
```

### Step 10 – Practice the Pipe

Use the `%>%` operator to pass each result to the next tidyverse verb.

```{r step_10, eval=FALSE}
m2 = data %>% lm(formula = log(cost) ~ unit)
m2
```

### Step 11 – Create `data4`

ln(cost) = 8.63484 ln$ + 0.03726 ln$ * unit.

```{r step_11, eval=FALSE}
data4 = tibble(
  unit = 1:150,
  pred = predict(m2, newdata = tibble(unit))
) %>%
  # remember to exponentiate the prediction!
  mutate(pred = exp(pred))
```

### Step 12 – Start a ggplot

And visualize. Initialize a ggplot so you can layer geoms and customise aesthetics.

```{r step_12, eval=FALSE}
ggplot() +
  geom_point(data = data, mapping = aes(x = unit, y = cost)) +
  geom_line(data = data3, mapping = aes(x = unit, y = pred), color = "blue") +
  geom_line(data = data4, mapping = aes(x = unit, y = pred), color = "red") +
  geom_point(data = data2, mapping = aes(x = unit, y = pred), color = "blue")
```

### Step 13 – Clear Objects

Cleanup. Remove objects from the environment to prevent name clashes.

```{r step_13, eval=FALSE}
rm(list = ls())
```

## Learning Checks

**Learning Check 1.** How do you run the entire workshop script after you have stepped through each section interactively?

<details>
<summary>Show answer</summary>

Use `source(file.path("workshops", "09_architecting_cost_function_examples.R"))` from the Console or press the Source button while the script is active.

</details>

**Learning Check 2.** Why does the script begin by installing or loading packages before exploring the exercises?

<details>
<summary>Show answer</summary>

Those commands make sure the required libraries are available so every subsequent code chunk runs without missing-function errors.

</details>

**Learning Check 3.** How does the `%>%` pipeline help you reason about multi-step transformations in this script?

<details>
<summary>Show answer</summary>

It keeps each operation in sequence without creating temporary variables, so you can narrate the data story line by line.

</details>

**Learning Check 4.** How can you build confidence that a newly defined function behaves as intended?

<details>
<summary>Show answer</summary>

Call it with the sample input from the script, examine the output, then try a new input to see how the behaviour changes.

</details>

**Learning Check 5.** What experiment can you run on the `ggplot` layers to understand how aesthetics map to data?

<details>
<summary>Show answer</summary>

Switch one aesthetic (for example `color` to `fill` or tweak the geometry) and re-run the chunk to observe the difference.

</details>
