---
title: "[14] Enumeration tutorial foundations Guide"
output:
  md_document:
    variant: gfm
output_dir: ../workshops
knitr:
  opts_knit:
    root.dir: ..
---

This tutorial complements `14_enumeration_foundations.R` and unpacks the workshop on enumeration tutorial foundations. You will see how it advances the Enumeration sequence while building confidence with base R and tidyverse tooling.

## Setup

- Ensure you have opened the `archr` project root (or set your working directory there) before running any code.
- Open the workshop script in RStudio so you can execute lines interactively with `Ctrl+Enter` or `Cmd+Enter`.
- Create a fresh R session to avoid conflicts with leftover objects from earlier workshops.

## Skills

- Navigate the script `14_enumeration_foundations.R` within the Enumeration module.
- Connect the topic "Enumeration tutorial foundations" to systems architecting decisions.
- Load packages with `library()` and verify they attach without warnings.
- Import data files with `readr` helpers and inspect the resulting objects.
- Chain tidyverse verbs with `%>%` to explore stakeholder or architecture tables.

## Application

### Step 1 – Load Packages

Load packages. Attach dplyr to make its functions available.

```{r step_01, eval=FALSE}
library(dplyr) # data wrangling
library(readr) # read_csv etc.
library(tidyr) # expand_grid(), etc.
library(archr) # enumerate_binary(), etc.
```

### Step 2 – Run the Code Block

How would we make binary decisions?

```{r step_02, eval=FALSE}
enumerate_binary(n = 2)
enumerate_binary(n = 2, .id = TRUE)
enumerate_binary(n = 4)
enumerate_binary(n = 16)
```

### Step 3 – Practice the Pipe

How many rows (architectures) in a 16-binary-decision architectural space?

```{r step_03, eval=FALSE}
enumerate_binary(n = 16) %>% nrow()
```

### Step 4 – Run the Code Block

What about standard form decisions?

```{r step_04, eval=FALSE}
enumerate_sf(n = c(2, 3))
enumerate_sf(n = c(2, 3, 4))
enumerate_sf(n = c(2, 3, 4), .id = TRUE)
```

### Step 5 – Run the Code Block

How about downselecting? n is the number of choices, k is the MAX number of items you can select.

```{r step_05, eval=FALSE}
enumerate_ds(n = 2, k = 1)
enumerate_ds(n = 2, k = 1)
enumerate_ds(n = 5, k = 3)
```

### Step 6 – Run the Code Block

Want to check the documentation? Use ?function_name or ?package_name::function_name It will show up in the 'Help' menu Works for ANY package. Most of archr's functions have at least some documentation.

```{r step_06, eval=FALSE}
?enumerate_ds
# Try it for mutate!
?dplyr::mutate
```

### Step 7 – Run the Code Block

Can we replicate enumerate_binary() using dplyr and tidyr functions? Yes! expand_grid() is a magical function that builds a grid of every possible combination of the items you supply it.

```{r step_07, eval=FALSE}
expand_grid(
  tibble(d1 = c(0, 1)),
  tibble(d2 = c(0, 1))
)
# Compare with enumerate_binary() - it's the same!
enumerate_binary(n = 2)
```

### Step 8 – Run the Code Block

Can work for standard form too.

```{r step_08, eval=FALSE}
expand_grid(
  tibble(d1 = c(0, 1)),
  tibble(d2 = c(0, 1, 2))
)
# Check it!
enumerate_sf(n = c(2,3))
```

### Step 9 – Run the Code Block

Can also do creative combinations like this, which replicates the d2-d3 data.frame for each item in the d1 data.frame.

```{r step_09, eval=FALSE}
expand_grid(
  tibble(d1 = c(0, 1)),
  tibble(d2 = c(0, 0, 0),
         d3 = c(0, 2, 3))
)
```

### Step 10 – Run the Code Block

More flexible usage! Execute the block and pay attention to the output it produces.

```{r step_10, eval=FALSE}
expand_grid(
  enumerate_binary(n = 2),
  # You'll need to 'rename' your columns to avoid errors
  enumerate_sf(n = c(2,3)) %>% select(d3 = 1, d4 = 2)
)
```

## Learning Checks

**Learning Check 1.** How do you run the entire workshop script after you have stepped through each section interactively?

<details>
<summary>Show answer</summary>

Use `source(file.path("workshops", "14_enumeration_foundations.R"))` from the Console or press the Source button while the script is active.

</details>

**Learning Check 2.** Why does the script begin by installing or loading packages before exploring the exercises?

<details>
<summary>Show answer</summary>

Those commands make sure the required libraries are available so every subsequent code chunk runs without missing-function errors.

</details>

**Learning Check 3.** When you import data in this workshop, what should you inspect right after the read call?

<details>
<summary>Show answer</summary>

Check the tibble in the Environment pane (or print it) to confirm column names and types look correct.

</details>

**Learning Check 4.** How does the `%>%` pipeline help you reason about multi-step transformations in this script?

<details>
<summary>Show answer</summary>

It keeps each operation in sequence without creating temporary variables, so you can narrate the data story line by line.

</details>
