---
title: "[20] Custom metric development Guide"
output:
  md_document:
    variant: gfm
output_dir: ../workshops
knitr:
  opts_knit:
    root.dir: ..
---

This tutorial complements `20_enumeration_custom_metrics.R` and unpacks the workshop on custom metric development. You will see how it advances the Enumeration sequence while building confidence with base R and tidyverse tooling.

## Setup

- Ensure you have opened the `archr` project root (or set your working directory there) before running any code.
- Open the workshop script in RStudio so you can execute lines interactively with `Ctrl+Enter` or `Cmd+Enter`.
- Create a fresh R session to avoid conflicts with leftover objects from earlier workshops.

## Skills

- Navigate the script `20_enumeration_custom_metrics.R` within the Enumeration module.
- Connect the topic "Custom metric development" to systems architecting decisions.
- Load packages with `library()` and verify they attach without warnings.
- Chain tidyverse verbs with `%>%` to explore stakeholder or architecture tables.
- Define custom functions to package repeatable logic.

## Application

### Step 1 – Load Packages

Attach dplyr to make its functions available.

```{r step_01, eval=FALSE}
library(dplyr)
library(archr)
```

### Step 2 – Define `get_percentages()`

Can be hard to compare visually, but let's check them numerically...

```{r step_02, eval=FALSE}
get_percentages = function(x, did = "d2"){
  # Testing values
  #x = mini$d2
```

### Step 3 – Practice the Pipe

Get the tally per alternative.

```{r step_03, eval=FALSE}
  tally = tibble(altid = x) %>%
    group_by(altid) %>%
    summarize(count = n())
  # Calculate percentages
  percentages = tally %>% 
    mutate(total = sum(count)) %>%
    mutate(percent = count / total) %>%
    mutate(label = round(percent * 100, 1)) %>%
    mutate(label = paste0(label, "%"))
  # Add a decision id and format
  percentages = percentages %>% 
    mutate(did = did) %>%
    select(did, altid, count, total, percent, label)
```

### Step 4 – Run the Code Block

Execute the block and pay attention to the output it produces.

```{r step_04, eval=FALSE}
  return(percentages)
}
```

### Step 5 – Create `a`

Make an architecture matrix of 3 binary decisions.

```{r step_05, eval=FALSE}
a = enumerate_sf(n = c(2,2,2))
```

### Step 6 – Practice the Pipe

Repeat this code several times.

```{r step_06, eval=FALSE}
sample = a %>%
  group_by(d1) %>%
  sample_n(size = 2)
```

### Step 7 – Run the Code Block

Execute the block and pay attention to the output it produces.

```{r step_07, eval=FALSE}
get_percentages(x = sample$d1, did = "d1")
get_percentages(x = sample$d2, did = "d2")
get_percentages(x = sample$d3, did = "d3")
```

### Step 8 – Practice the Pipe

One way to improve balance is to stratify by MULTIPLE variables. (though NEVER by ALL variables).

```{r step_08, eval=FALSE}
sample = a %>%
  group_by(d1, d2) %>%
  sample_n(size = 2)
```

### Step 9 – Run the Code Block

Execute the block and pay attention to the output it produces.

```{r step_09, eval=FALSE}
get_percentages(x = sample$d1, did = "d1")
get_percentages(x = sample$d2, did = "d2")
get_percentages(x = sample$d3, did = "d3")
```

### Step 10 – Clear Objects

A little better! Remove objects from the environment to prevent name clashes.

```{r step_10, eval=FALSE}
rm(list = ls())
```

## Learning Checks

**Learning Check 1.** How do you run the entire workshop script after you have stepped through each section interactively?

<details>
<summary>Show answer</summary>

Use `source(file.path("workshops", "20_enumeration_custom_metrics.R"))` from the Console or press the Source button while the script is active.

</details>

**Learning Check 2.** Why does the script begin by installing or loading packages before exploring the exercises?

<details>
<summary>Show answer</summary>

Those commands make sure the required libraries are available so every subsequent code chunk runs without missing-function errors.

</details>

**Learning Check 3.** How does the `%>%` pipeline help you reason about multi-step transformations in this script?

<details>
<summary>Show answer</summary>

It keeps each operation in sequence without creating temporary variables, so you can narrate the data story line by line.

</details>

**Learning Check 4.** How can you build confidence that a newly defined function behaves as intended?

<details>
<summary>Show answer</summary>

Call it with the sample input from the script, examine the output, then try a new input to see how the behaviour changes.

</details>
