---
title: "[28] Pareto fronts exploration Guide"
output:
  md_document:
    variant: gfm
output_dir: ../workshops
knitr:
  opts_knit:
    root.dir: ..
---

This tutorial complements `28_evaluation_pareto_fronts.R` and unpacks the workshop on pareto fronts exploration. You will see how it advances the Evaluation sequence while building confidence with base R and tidyverse tooling.

## Setup

- Ensure you have opened the `archr` project root (or set your working directory there) before running any code.
- Open the workshop script in RStudio so you can execute lines interactively with `Ctrl+Enter` or `Cmd+Enter`.
- Create a fresh R session to avoid conflicts with leftover objects from earlier workshops.

## Skills

- Navigate the script `28_evaluation_pareto_fronts.R` within the Evaluation module.
- Connect the topic "Pareto fronts exploration" to systems architecting decisions.
- Load packages with `library()` and verify they attach without warnings.
- Chain tidyverse verbs with `%>%` to explore stakeholder or architecture tables.
- Iterate on visualisations built with `ggplot2`.

## Application

### Step 1 – Load Packages

SETUP ################################# Load packages.

```{r step_01, eval=FALSE}
library(dplyr)
library(archr)
library(ggplot2)
# Load data
data("donuts", package = "archr")
donuts
```

### Step 2 – Create `m`

PARETO FRONT ############################ Make a table m we'll build Enumerate architectures, including an .id.

```{r step_02, eval=FALSE}
m = enumerate_binary(n = nrow(donuts), .id = TRUE)
```

### Step 3 – Practice the Pipe

Record the architectures as a matrix.

```{r step_03, eval=FALSE}
a = m %>% select(-id) %>% as.matrix()
```

### Step 4 – Practice the Pipe

Add in columns for total metrics.

```{r step_04, eval=FALSE}
m = m %>% 
  mutate(
    # For this example, we'll calculate benefit and cost with matrix multiplication,
    # but you could calculate cost or benefit many different ways
    # Matrix multiple the matrix by the vector of benefits
    benefit = a %*% donuts$benefit,
    # Matrix multiple the matrix by the vector of costs
    cost =   a %*% donuts$cost)
```

### Step 5 – Practice the Pipe

View last five. Use the `%>%` operator to pass each result to the next tidyverse verb.

```{r step_05, eval=FALSE}
m %>% select(id, benefit, cost) %>% tail(5)
```

### Step 6 – Start a ggplot

Initialize a ggplot so you can layer geoms and customise aesthetics.

```{r step_06, eval=FALSE}
ggplot() +
  geom_point(data = m, mapping = aes(x = benefit, y = cost))
```

### Step 7 – Practice the Pipe

Get the ids of the pareto front with pareto().

```{r step_07, eval=FALSE}
m = m %>%
  mutate(front = pareto(x = cost, y = -benefit))
```

### Step 8 – Practice the Pipe

Use the `%>%` operator to pass each result to the next tidyverse verb.

```{r step_08, eval=FALSE}
m %>% select(id, cost, benefit, front)
# ?archr::pareto()
```

### Step 9 – Start a ggplot

Initialize a ggplot so you can layer geoms and customise aesthetics.

```{r step_09, eval=FALSE}
ggplot() +
  geom_point(data = m, mapping = aes(x = benefit, y = cost),
             color = "grey") +
  geom_line(data = m %>% filter(front == TRUE),
            mapping = aes(x = benefit, y = cost),
            color = "red")
```

### Step 10 – Practice the Pipe

Thresholds ##################################### Show just a slice of the architectures based on thresholds.

```{r step_10, eval=FALSE}
m %>% 
  filter(front == TRUE & benefit > 20 & cost < 100)
```

### Step 11 – Start a ggplot

Add those thresholds for context usinge geom_hline and geom_vline!

```{r step_11, eval=FALSE}
ggplot() +
  geom_point(data = m, mapping = aes(x = benefit, y = cost),
             color = "grey") +
  geom_line(data = m %>% filter(front == TRUE),
            mapping = aes(x = benefit, y = cost),
            color = "red") +
  geom_hline(yintercept = 100, color = "blue") +
  geom_vline(xintercept = 20, color = "blue")
```

### Step 12 – Run the Code Block

Execute the block and pay attention to the output it produces.

```{r step_12, eval=FALSE}
m
```

### Step 13 – Clear Objects

Cleanup ! Remove objects from the environment to prevent name clashes.

```{r step_13, eval=FALSE}
rm(list = ls())
```

## Learning Checks

**Learning Check 1.** How do you run the entire workshop script after you have stepped through each section interactively?

<details>
<summary>Show answer</summary>

Use `source(file.path("workshops", "28_evaluation_pareto_fronts.R"))` from the Console or press the Source button while the script is active.

</details>

**Learning Check 2.** Why does the script begin by installing or loading packages before exploring the exercises?

<details>
<summary>Show answer</summary>

Those commands make sure the required libraries are available so every subsequent code chunk runs without missing-function errors.

</details>

**Learning Check 3.** How does the `%>%` pipeline help you reason about multi-step transformations in this script?

<details>
<summary>Show answer</summary>

It keeps each operation in sequence without creating temporary variables, so you can narrate the data story line by line.

</details>

**Learning Check 4.** What experiment can you run on the `ggplot` layers to understand how aesthetics map to data?

<details>
<summary>Show answer</summary>

Switch one aesthetic (for example `color` to `fill` or tweak the geometry) and re-run the chunk to observe the difference.

</details>
