---
title: "[41] Sensitivity and connectivity lab Guide"
output:
  md_document:
    variant: gfm
output_dir: ../workshops
knitr:
  opts_knit:
    root.dir: ..
---

This tutorial complements `41_evaluation_sensitivity_connectivity_lab.R` and unpacks the workshop on sensitivity and connectivity lab. You will see how it advances the Evaluation sequence while building confidence with base R and tidyverse tooling.

## Setup

- Ensure you have opened the `archr` project root (or set your working directory there) before running any code.
- Open the workshop script in RStudio so you can execute lines interactively with `Ctrl+Enter` or `Cmd+Enter`.
- Create a fresh R session to avoid conflicts with leftover objects from earlier workshops.

## Skills

- Navigate the script `41_evaluation_sensitivity_connectivity_lab.R` within the Evaluation module.
- Connect the topic "Sensitivity and connectivity lab" to systems architecting decisions.
- Load packages with `library()` and verify they attach without warnings.
- Chain tidyverse verbs with `%>%` to explore stakeholder or architecture tables.
- Iterate on visualisations built with `ggplot2`.

## Application

### Step 1 – Load Packages

Attach dplyr to make its functions available.

```{r step_01, eval=FALSE}
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(archr)
```

### Step 2 – Run the Code Block

Execute the block and pay attention to the output it produces.

```{r step_02, eval=FALSE}
source("workshops/00_sensitivity_connectivity_utilities.R")
```

### Step 3 – Practice the Pipe

Replicate the exact data structure from the slides.

```{r step_03, eval=FALSE}
data = enumerate_sf(n = c(2, 3, 2)) %>%
  # Add metrics
  mutate(m1 = c(33,73, 40, 80, 30, 70, 66, 146, 80, 160, 60, 140),
         m2 = c(0.8910, 0.9801, 0.8991, 0.9890, 0.4500, 0.4950, 0.9899, 0.9998, 0.9900, 0.9999, 0.7425, 0.7499)) %>%
  mutate(d1 = d1 + 1, d2 = d2 + 1, d3 = d3+1) %>%
  select(d1 = d3, d2 = d2, d3 = d1, m1, m2)
```

### Step 4 – Run the Code Block

Execute the block and pay attention to the output it produces.

```{r step_04, eval=FALSE}
data
```

### Step 5 – Run the Code Block

Execute the block and pay attention to the output it produces.

```{r step_05, eval=FALSE}
connectivity(data = data, decision_i = "d2", decisions = c("d1","d2","d3"), metric = "m1")
```

### Step 6 – Run the Code Block

Execute the block and pay attention to the output it produces.

```{r step_06, eval=FALSE}
sensitivity(data = data,decision_i = "d2", metric = "m1")
```

### Step 7 – Create `points`

VISUAL ###################################. Create the object `points` so you can reuse it in later steps.

```{r step_07, eval=FALSE}
points = expand_grid(
  decision = c("d1","d2","d3"),
  metric = c("m1","m2")
) %>%
  group_by(decision, metric) %>%
  #rowwise() %>% 
  summarize(
    c = connectivity(data = data, decision_i = decision, decisions = c("d1","d2","d3"), metric = metric),
    s = sensitivity(data = data, decision_i = decision, metric = metric),
    .groups = "drop"
  )
```

### Step 8 – Run the Code Block

Execute the block and pay attention to the output it produces.

```{r step_08, eval=FALSE}
points
```

### Step 9 – Start a ggplot

Initialize a ggplot so you can layer geoms and customise aesthetics.

```{r step_09, eval=FALSE}
ggplot() +
  geom_point(data = points, mapping = aes(
    x = c, y = s, color = decision),
    size = 10) +
  geom_text(data = points, mapping = aes(
    x = c, y = s, label = decision), 
    color = "white") +
  facet_wrap(~metric, scales = "free") +
  theme(legend.position = "none")
```

### Step 10 – Practice the Pipe

Use the `%>%` operator to pass each result to the next tidyverse verb.

```{r step_10, eval=FALSE}
pointsm1 = points %>%
  filter(metric == "m1")
```

### Step 11 – Start a ggplot

Initialize a ggplot so you can layer geoms and customise aesthetics.

```{r step_11, eval=FALSE}
ggplot() +
  geom_point(data = pointsm1, mapping = aes(
    x = c, y = s, color = decision),
    size = 10) +
  geom_text(data = pointsm1, mapping = aes(
    x = c, y = s, label = decision), 
    color = "white") +
  theme(legend.position = "none") +
  geom_vline(xintercept = 12.5) +
  geom_hline(yintercept = 30) +
  scale_x_continuous(limits = c(0, 25)) +
  scale_y_continuous(limits = c(0, 60)) +
  labs(x = "Connectivity Score (Metric X)",
       y = "Sensitivity Score (Metric X)")
```

## Learning Checks

**Learning Check 1.** How do you run the entire workshop script after you have stepped through each section interactively?

<details>
<summary>Show answer</summary>

Use `source(file.path("workshops", "41_evaluation_sensitivity_connectivity_lab.R"))` from the Console or press the Source button while the script is active.

</details>

**Learning Check 2.** Why does the script begin by installing or loading packages before exploring the exercises?

<details>
<summary>Show answer</summary>

Those commands make sure the required libraries are available so every subsequent code chunk runs without missing-function errors.

</details>

**Learning Check 3.** How does the `%>%` pipeline help you reason about multi-step transformations in this script?

<details>
<summary>Show answer</summary>

It keeps each operation in sequence without creating temporary variables, so you can narrate the data story line by line.

</details>

**Learning Check 4.** What experiment can you run on the `ggplot` layers to understand how aesthetics map to data?

<details>
<summary>Show answer</summary>

Switch one aesthetic (for example `color` to `fill` or tweak the geometry) and re-run the chunk to observe the difference.

</details>
